---
title: "p8105_hw5_xl3214"
author: "Xuan Lu"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This analysis delves into the grim reality of unsolved homicides across various cities in the United States. Drawing inspiration from a [2018 Washington Post investigation](https://www.washingtonpost.com/graphics/2018/investigations/where-murders-go-unsolved/) which highlighted the concerning rates of murders going unsolved, particularly in certain urban areas, this report seeks to illuminate patterns in unsolved homicide cases. Utilizing a comprehensive dataset detailing homicides, this analysis not only identifies cities with high proportions of unsolved cases but also examines the demographic and geographical nuances underlying these statistics. The rationale behind this study stems from the urgent need to address and understand the factors contributing to the lack of resolution in these cases, as this has profound implications for justice and public safety.

## Data Analysis

```{r Load necessary packages, message=FALSE, results='hide', warning=FALSE}
library(tidyverse)
library(dplyr)
library(broom)
library(ggplot2)
```


```{r Load dataset, message=FALSE, warning=FALSE}
homicide_raw <- read.csv(file = "data_files/homicide-data.csv")
```

The raw data has `r ncol(homicide_raw)` columns and `r nrow(homicide_raw)` rows. It contains variables *`r colnames(homicide_raw)`*. 

```{r}
homicide <- homicide_raw |>
  janitor::clean_names() |>
  mutate(city_state = paste(city, state, sep = ", ")) |>
  mutate(victim_race = as.factor(victim_race)) |>
  mutate(victim_sex = as.factor(victim_sex)) |>
  mutate(victim_age = as.factor(victim_age)) |> 
  mutate(disposition = as.factor(disposition))
summary(homicide)
```

```{r city_state and disposition, message=FALSE, warning=FALSE, fig.height=10, fig.width=8}
homicide_sum <- homicide |> 
  group_by(city_state, disposition) |>
  summarise(total_homicides = n())
homicide_sum |> 
  pivot_wider(names_from = disposition, values_from = total_homicides) |>
  knitr::kable()

proportion_test_by_city <- function(data, city_column, disposition_column) {
  data |>
    group_by({{ city_column }}) |>
    summarize(
      unsolved_count = sum({{ disposition_column }} %in% c("Closed without arrest", "Open/No arrest"), na.rm = TRUE),
      total_cases = n(),
      prop_test_result = list(prop.test(unsolved_count, total_cases, correct = FALSE))
    ) |>
    mutate(tidy_result = map(prop_test_result, broom::tidy)) |>
    select(-prop_test_result) |>
    unnest(tidy_result) |>
    select(city = {{ city_column }}, estimate, conf.low, conf.high)
}

homicide_summary <- proportion_test_by_city(homicide, city_state, disposition)

# Sort cities by the estimated proportion of unsolved homicides
homicide_summary_sorted <- homicide_summary |>
  arrange(desc(estimate))

# Create the plot
ggplot(homicide_summary_sorted, aes(x = reorder(city, estimate), y = estimate)) +
  geom_point() +  # Add points for estimates
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +  # Add error bars
  coord_flip() +  # Flip coordinates to make it easier to read city names
  labs(
    title = "Proportion of Unsolved Homicides by City",
    x = "City",
    y = "Estimated Proportion of Unsolved Homicides"
  ) +
  theme_minimal()
```

## Findings and Interpretation

The analysis revealed significant variations in the proportion of unsolved homicides across different cities. Notably, cities like Baltimore, Detroit, and Chicago emerged with a higher percentage of unsolved cases, aligning with national trends reported in major investigations. This pattern raises critical questions about law enforcement efficacy, resource allocation, and possible systemic issues within these urban centers.

The data also brought to light the demographic aspects of the victims, showing disparities in the resolution of cases based on factors like race and age. Such disparities point towards potential underlying biases or systemic challenges in the criminal justice system that need addressing.